---
title: "CDC2021_Statistical_Analysis"
author: "Feng Zhang"
date: "9/30/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Clean Data of New York 
We decided to examine the clean data of New York's counties. In this set of commands, we are trying to figure out the % of insured and uninsured folks. 

```{r}
#To set the directory path for R to get the data
setwd("C:/Users/tommy/Downloads")

New_York_Clean <- read.csv("cleanedNY.csv")
New_York_Clean_Percentage <- read.csv("cleanedNYpercentages.csv")

head(New_York_Clean)
head(New_York_Clean_Percentage)

```

#To see what are the values in NY data and data types
```{r}
str(New_York_Clean)
str(New_York_Clean, list.len = 4)

lapply(New_York_Clean, typeof)
```


# Summary Statistics of Uninsured people in New York
```{r}
New_York_uninsured = New_York_Clean %>% select (State, County, Uninsured.Population..Excluding.Undocumented.)

head(New_York_uninsured)

Uninsured_population <-as.numeric(gsub(",","", New_York_uninsured[,3] ))
summary(Uninsured_population, title = "Uninsured population in the state of New York")

```
# Proportion of race and income 
Comparing the race and income of the uninsured population of NY
Then doing a simple linear regression to find if the two variables have a linear relationship 
```{r}
# Extract the columns for race & Income levels
# convert characters to #s
# Then do a simple regression or correlation to see if the the two have a relationship
Race = New_York_Clean %>% select (State,County,Spanish.Hispanic.Latino.Origin, White.Non.Latino, Black.Non.Latino, Asian...Native.Hawaiian...Pac.Is, American.Indian...Alaska.Native, Multi.racial.or.Other)

head(Race)

State <- as.character(Race[,1])
County <- as.character(Race[,2])

Spanish_Origin <- as.numeric(gsub(",","", Race[,3]))
Spanish_Origin[is.na(Spanish_Origin)] <- 0

White_Non_Latino <- as.numeric(gsub(",","", Race[,4]))

Black_Non_Latino <- as.numeric(gsub(",","", Race[,5]))
Black_Non_Latino[is.na(Black_Non_Latino)] <- 0

AAPI <- as.numeric(gsub(",","", Race[,6]))
AAPI[is.na(AAPI)] <- 0

Native_American_Alaska_Native <- as.numeric(gsub(",","", Race[,7]))
Native_American_Alaska_Native[is.na(Native_American_Alaska_Native)] <- 0

Multi_Racial_Other<- as.numeric(gsub(",","", Race[,8]))
Multi_Racial_Other[is.na(Multi_Racial_Other)] <- 0
#New Table of Race info
Race_numbers = data.frame(Spanish_Origin, White_Non_Latino, Black_Non_Latino, AAPI, Native_American_Alaska_Native, Multi_Racial_Other)  
head(Race_numbers)


```

```{r}
Income = New_York_Clean %>% select(State,County,HIU.Income...100..FPL, HIU.Income.100.138..FPL, HIU.Income.139.249..FPL, HIU.Income.250.400..FPL, HIU.Income.400..FPL)

head(Income)

#HIU Income < 100% FPL
Income_less_100 <- as.numeric(gsub(",", "", Income[,3]))
Income_less_100[is.na(Income_less_100)] <- 0

#HIU Income 100-138% FPL
Income_138 <- as.numeric(gsub(",", "", Income[,4]))
Income_138[is.na(Income_138)] <- 0

#HIU Income 139-249% FPL
Income_249 <- as.numeric(gsub(",", "", Income[,5]))
Income_249[is.na(Income_249)] <- 0

#HIU Income 250-400% FPL
Income_400 <- as.numeric(gsub(",", "", Income[,6]))
Income_400[is.na(Income_400)] <- 0

#HIU Income 400% FPL
Income_400_more <- as.numeric(gsub(",", "", Income[,7]))
Income_400_more[is.na(Income_400_more)] <- 0


Income_numbers = data.frame(Income_less_100 ,Income_138,Income_249, Income_400, Income_400_more)

head(Income_numbers)


```
```{r}
# Now For Correlation
correlation <-cor(Race_numbers, Income_numbers)


plot(correlation, xlab = "Racial Groups ", ylab = "Income Levels", main = "Racial groups vs. Income levels")

```

#Proportion of Education Level

```{r}
Education = New_York_Clean %>% select(State, County, Less.than.High.School, High.School.Diploma, College.Grad)

head(Education)

Less_High_School <-as.numeric(gsub(",", "", Education[,3]))
Less_High_School [is.na(Less_High_School )] <- 0

High_School <-as.numeric(gsub(",", "", Education[,4]))
High_School[is.na(High_School)] <- 0

College_Grad <-as.numeric(gsub(",", "", Education[,5]))
College_Grad[is.na(College_Grad)] <- 0

Education_Numbers = data.frame(Less_High_School, High_School, College_Grad)
head(Education_Numbers)
```
# Correlation and Linear Regression
Assumptions for the Linear Regression

Linearity: The relationship between X and the mean of Y is linear.
Homoscedasticity: The variance of residual is the same for any value of X.
Independence: Observations are independent of each other.
Normality: For any fixed value of X, Y is normally distributed.

```{r}
x1 <- Uninsured_population
y1 <- Less_High_School

lm1 <- lm(y1 ~ x1)

plot(x1, y1, main = "Uninsured vs. less than high school education", xlab = "Uninsured Population", ylab = "Less Than High School Education")
abline(a = -83.6019 , b = 0.2031 , col = "red")
# Summary of the linear regression between uninsured individuals v. less than high school education
summary(lm1)

x2 <- Uninsured_population
y2 <- High_School
lm2 <- lm(y2 ~ x2)
plot(x2, y2, main = "Uninsured vs. High School Diploma", xlab = "Uninsured Population", ylab = "High School Diploma")
abline(a =  454.700 , b = 0.544 , col = "red")
# Summary of the linear regression between uninsured individuals v. high school education
summary(lm2)

x3 <- Uninsured_population
y3 <- College_Grad
lm3 <- lm(y3 ~ x3)
plot(x3, y3, main = "Uninsured vs. College Graduate", xlab = "Uninsured Population", ylab = "College Graduate")
abline(a =   -388.4040  , b = 0.2531 , col = "red")
#Summary of the linear regression between uninsured individuals v. College/ college graduate education
summary(lm3)



```